### 2013-08-12 doc/dissertation/document/chapters/spn_equations/spn_equations.tex
updating dissertation build

In this chapter, we briefly derive the simplified $P_N$ ($SP_N$)
equations, closely following the work of Evans
\cite{evans_simplified_2013}\footnote{The $SP_N$ derivations in this
  chapter and the associated appendices are heavily based on those
  presented by Evans in \cite{evans_simplified_2013}}, in order to gain
full understanding of the underlying system and its potential behavior
in a Monte Carlo context. From the $P_N$ equations, we apply a set of
approximations to yield the $SP_N$ equations for fixed source and
criticality problems. Using the fully-formed linear operator for the
transport problem, we explore solutions to the $SP_N$ equations with
Monte Carlo Synthetic Acceleration using a challenging light water
reactor fuel assembly criticality calculation as the driving
problem. Several difficulties arise when applying MCSA to these
problems that were not observed when investigating the simpler
transport systems.

@@

In addition to the significantly larger number of histories required
to achieve convergence for ill-conditioned problems another penalty is
paid due to histories that take longer to
compute. Figure~\ref{fig:breakdown_time} gives the CPU time in seconds
required to compute a single random walk averaged over the entire set
of histories run in the calculation over all iterations. As the
spectral radius increases (correlating to a higher ratio of scattering
in the system) the random walk lengths increase, using more CPU time
to finish the computation. Compared to spectral radii of 0.5, larger
spectral radii over 0.97 have histories that require two orders of
magnitude more computation time. This significant increase in
computation time per history coupled with the significant increase in
the number of histories required to converge is evidence that for
problems with spectral radii above $\approx 0.97$, using MCSA to solve
any problems of interest is entirely ineffective and not practical. We
therefore require a more expansive set of preconditioning techniques
to move the eigenvalue spectrum of the $SP_N$ problem into a regime in
which MCSA is more applicable and in which performance is improved.
\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=4.25in]{chapters/spn_equations/breakdown_time.pdf}
  \end{center}
  \caption{\textbf{CPU time per history as a function of spectral
      radius for the neutron diffusion problem.} \textit{As the
      spectral radius grows, so do the length of the random
      walks. Longer random walks require more CPU time to compute.}}
  \label{fig:breakdown_time}
\end{figure}

>>>>

In this chapter, we briefly derive the simplified $P_N$ ($SP_N$)
equations, closely following the work of Evans
\cite{evans_simplified_2013}, in order to gain full understanding of
the underlying system and its potential behavior in a Monte Carlo
context. From the $P_N$ equations, we apply a set of approximations to
yield the $SP_N$ equations for fixed source and criticality
problems. Using the fully-formed linear operator for the transport
problem, we explore solutions to the $SP_N$ equations with Monte Carlo
Synthetic Acceleration using a challenging light water reactor fuel
assembly criticality calculation as the driving problem. Several
difficulties arise when applying MCSA to these problems that were not
observed when investigating the simple transport system in the
previous chapter.

@@

In addition to the significantly larger number of histories required
to achieve convergence for ill-conditioned problems another penalty is
paid due to histories that take longer to
compute. Figure~\ref{fig:breakdown_time} gives the CPU time in seconds
required to compute a single random walk averaged over the entire set
of histories run in the calculation over all iterations. As the
spectral radius increases (correlating to a higher ratio of scattering
in the system) the random walk lengths increase, using more CPU time
to finish the computation. Compared to spectral radii of 0.5, larger
spectral radii over 0.97 have histories that require two orders of
magnitude more computation time. This significant increase in
computation time per history coupled with the significant increase in
the number of histories required to converge is evidence that for
problems with spectral radii above $\approx 0.97$, using MCSA to solve
any problems of interest is entirely ineffective and not practical. At
iteration matrix spectral radii this large, using significant numbers
of stochastic histories in the Monte Carlo solve are not enough to
reduce the stochastic uncertainty in the correction, causing the MCSA
sequence to diverge instead of accelerating convergence. We therefore
require a more expansive set of preconditioning techniques to move the
eigenvalue spectrum of the $SP_N$ problem into a regime in which MCSA
is more applicable and in which performance is improved.
\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=4.25in]{chapters/spn_equations/breakdown_time.pdf}
  \end{center}
  \caption{\textbf{CPU time per history as a function of spectral
      radius for the neutron diffusion problem.} \textit{As the
      spectral radius grows, so do the length of the random
      walks. Longer random walks require more CPU time to compute.}}
  \label{fig:breakdown_time}
\end{figure}
